---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
author: André Tannús
---

## Libraries.

Some effort was put into making this document dependant on a few libraries as possible.
The only requirement is the ggplot2 library.
```{r}
library("ggplot2")
```

## Loading and preprocessing the data

The data is loaded into an R data frame.
```{r}
data <- read.csv("activity.csv")
```

The date column is parsed as Date objects.
```{r}
data$date <- as.Date(data$date)
```

A copy of the complete cases is created.
```{r}
cc <- data[complete.cases(data),];
```


## What is mean total number of steps taken per day?

Aggregate the data adding the steps per day and plot a histogram.
```{r}
stepsDaily <- aggregate(list(steps=cc$steps), by=list(day=cc$date), FUN=sum)
hist(stepsDaily$steps, main = "Steps per day", xlab = "Steps in a day")
```

Mean number of steps taken per day:
```{r}
mean(stepsDaily$steps)
```

Median number of steps taken per day:
```{r}
median(stepsDaily$steps)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
